Modelling writing hesitations in text writing as finite mixture process

### 2023-07-17 report/results.Rmd
Add manuscript and simulation

```{r datasets, results='asis'}
tibble(Dataset = c("C2L1", "CATO", "SPL2", "PLanTra", "LIFT"),
                Source = c("@ronneberg2022process",
                              "@torrance2016adolescent",
                              "@torrance",
                              "@rossetti2022s",
                              "@nina_vandermeulen_2020_3893538"),
                Keylogger = c("EyeWrite", 
                           "EyeWrite", 
                           "CyWrite", 
                           "InputLog", 
                           "InputLog"),
                `Writing task` = c(
                                 "Argumentative essays", 
                                 "Expository texts",
                                 "Argumentative essays",
                                 "Text simplification",
                                 "Synthesis"),
                 `N (ppts)` = c(126, 26*2, 39, 47, 658),
#                n_texts = c(1, 2, 2, 2, NA),
#                n_sentences = c(),
#                n_words = c(),
                conditions = c(" ", 
                               "weak decoders / control; masked / unmasked",
                               "write in L1 / L2",
                               "pre / post test trained in plain language principles and control", 
                               "Various topics and genres"),
                `Mean age` = c(11.8, 16.9, 20.6, 23, 16.95),
                 Language = c( "Norwegian", 
                               "Norwegian", 
                               "English (L1) / Spanish (L2)", 
                               "English (L2)", 
                               "Dutch")) %>% 
  kable(caption = "Datasets in brief.") %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 12)

@@

  full_set <- d %>% 
    count(ppt, topic, genre) %>% 
    summarise(n = n(), .by = ppt) %>% 
    mutate(has_full_set = n == 4)

@@

  result[[4]] <- d %>% 
    filter(ppt %in% ppt_keep) %>% 
    select(ppt, enough_sentences) %>% 
    unique() %>% 
    count(enough_sentences)

  result[[1]] <- d %>%
    filter(!is.na(iki), 
           !is_edit,
           ppt %in% ppt_keep,
           enough_sentences) %>%
    select(ppt, iki, location) %>% 
    drop_na() %>% 
    mutate(too_fast = iki <= 50, 
           too_slow = iki >= 30000,
           across(ppt, ~as.numeric(factor(.)))) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean, 
                          se = se_bin)),
              .by = c(ppt)) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean), 
                     .names = "{.col}"))

  # Sample within each category random data points
  set.seed(365)
  result[[2]] <- d %>% 
    filter(!is.na(iki), 
           !is_edit,
           ppt %in% ppt_keep,
           enough_sentences,
           iki > 50,
           iki < 30000) %>% 
    mutate(keep = 1:n(),
           keep = sample(keep),
           keep = keep <= n_samples,
           .by = c(ppt, location, genre, topic)) %>% 
   summarise(across(keep, 
             list(mean = mean, 
                  se = se_bin)),
              .by = c(ppt, location)) %>% 
   summarise(across(starts_with("keep"), 
                     list(mean = mean), 
                     .names = "{.col}"), 
              .by = location)

   set.seed(365)
   result[[5]] <- d %>% 
    mutate(keep = 1:n(),
           keep = sample(keep),
           keep = keep <= n_samples, 
           .by = c(ppt, location, genre, topic)) %>% 
    filter(keep) %>% 
    mutate(ppt_keep = ppt %in% sample(unique(ppt), n_ppts)) %>% 
    select(ppt, ppt_keep) %>% 
    unique() %>% 
    count(ppt_keep)

@@

  result[[1]] <- d %>% 
    mutate(too_short = iki <= 50,
           too_fast = iki >= 30000,
           across(ppt, ~as.numeric(factor(.)))) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean, 
                          se = se_bin)),
              .by = c(ppt)) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean), 
                     .names = "{.col}"))

@@

  set.seed(365)
  result[[2]] <- d %>% 
    mutate(keep = 1:n(),
           keep = sample(keep),
           keep = keep <= n_samples,
           .by = c(ppt, location, Lang)) %>% 
   summarise(across(keep, 
             list(mean = mean, 
                  se = se_bin)),
              .by = c(ppt, location)) %>% 
   summarise(across(starts_with("keep"), 
                     list(mean = mean), 
                     .names = "{.col}"), 
              .by = location)

  return(result)
}
```

```{r}
# Get info about data reduction
n_samples <- 100
c2l1 <- get_c2l1("../data/c2l1.csv", n_samples = n_samples)
cato <- get_cato("../data/cato.csv", n_samples = n_samples)
lift <- get_lift("../data/lift.csv", n_samples = 50, n_ppts = 100)
plantra <- get_plantra("../data/plantra.csv", n_samples = n_samples)
spl2 <- get_spl2("../data/spl2.csv", n_samples = n_samples)

```{r eval = F}
Data reduction using random samples of subsets of participants for larger data sets

Also for larger data sets, only used participants that provided a min number of sentences

Removed transitions that were followed by an editing operation

Extremely short and long intervals

Add a description with general data filtering and a table with how many data were removed from the analysis by study.

Add final sample number of ppt, and transitions by location

@@

# Posterior by data set

## GUNNEXP2

### Fit to data

### Unconstrained mixture model

#### Posterior parameter estimates

@@

#### Masking effect

@@

### Cononstrained mixture model

#### Posterior parameter estimates

@@

## C2L1

The C2L1 data set comprises data Norwegian 6th graders--*N*=126, mean age 11 years 10 months--published in @ronneberg2022process. The children composed argumentative essays in Norwegian, a language with a relatively shallow orthography.

```{r eval = F}
d <- read_csv("../data/c2l1.csv") %>%
  filter(!is.na(iki), 
         iki > 50, 
         iki < 30000) %>%
  mutate(ppt = as.numeric(factor(subno)),
         condition = factor(str_c(location, sep = "_")),
         cond_num = as.integer(condition)) %>% 
  select(ppt, iki, condition, cond_num)

count(d, ppt) %>% 
  filter(n < 100)
pull(d, ppt) %>% unique() %>% length()
```

### Fit to data

```{r predictionsc2l1, fig.cap="C2L1 data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs.", eval = T}
file <- list.files("../stanout/c2l1", 
                   pattern = "sims", 
                   full.names = T)

plots <- get_pdens_plot(file, max_iki = 2000)

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Posterior parameter estimates of mixture model

@@

## CATO

### Fit to data

```{r predictionscato, fig.cap="CATO data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

file <- list.files("../stanout/cato", 
                    pattern = "sims", 
                    full.names = T)

plots <- get_pdens_plot(file, max_iki = 2000)

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Posterior parameter estimates of mixture model

@@

## PLanTra

### Methods

### Fit to data

```{r predictionsplantra, fig.cap="PLanTra data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

file <- list.files("../stanout/plantra", 
                   pattern = "sims", 
                   full.names = T)

plots <- get_pdens_plot(file, max_iki = 2000)

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Posterior parameter estimates of mixture model

@@

## LIFT

### Methods

### Fit to data

```{r predictionslift, fig.cap="LIFT data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

file <- list.files("../stanout/lift", 
                    pattern = "sims", 
                    full.names = T)

plots <- get_pdens_plot(file, max_iki = 2000)

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Posterior parameter estimates of mixture model

@@

## SPL2

### Methods

### Fit to data

```{r predictionsspl2, fig.cap="SPL2 data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs."}

file <- list.files("../stanout/spl2", pattern = "sims", full.names = T)

plots <- get_pdens_plot(file, max_iki = 2000)

grid.arrange(patchworkGrob(plots), 
             left = "Density",
             bottom = "Transition duration in msecs")
```

### Posterior parameter estimates of mixture model

@@

>>>>

```{r datasets, results='asis'}
tibble(Dataset = c("C2L1", "CATO", "SPL2", "PLanTra", "LIFT", "GUNNEXP2"),
                Source = c("@ronneberg2022process",
                              "@torrance2016adolescent",
                              "@torrance",
                              "@rossetti2022s",
                              "@nina_vandermeulen_2020_3893538", 
                           ""),
                Keylogger = c("EyeWrite", 
                           "EyeWrite", 
                           "CyWrite", 
                           "InputLog", 
                           "InputLog",
                           "EyeWrite"),
                `Writing task` = c(
                                 "Argumentative essays", 
                                 "Expository texts",
                                 "Argumentative essays",
                                 "Text simplification",
                                 "Synthesis", 
                                 ""),
                 `N (ppts)` = c(126, 26*2, 39, 47, 658, 45),
#                n_texts = c(1, 2, 2, 2, NA),
#                n_sentences = c(),
#                n_words = c(),
                conditions = c(" ", 
                               "weak decoders / control; masked / unmasked",
                               "write in L1 / L2",
                               "pre / post test trained in plain language principles and control", 
                               "Various topics and genres",
                               "masked / unmasked"),
                `Mean age` = c(11.8, 16.9, 20.6, 23, 16.95, NA),
                 Language = c( "Norwegian", 
                               "Norwegian", 
                               "English (L1) / Spanish (L2)", 
                               "English (L2)", 
                               "Dutch",
                               "Norwegian")) %>% 
  kable(caption = "Datasets in brief.") %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 12)

```
### GUNNEXP2

### C2L1

The C2L1 data set comprises data Norwegian 6th graders -- *N*=126, mean age 11 years 10 months -- published in @ronneberg2022process. The children composed argumentative essays in Norwegian, a language with a relatively shallow orthography.

### CATO

### PLanTra

### LIFT

### SPL2

@@

  result[[4]] <- d %>% 
    select(ppt, enough_sentences) %>% 
    unique() %>%
    filter(!enough_sentences) %>%
    drop_na() %>% 
    summarise(less_than_10_sentences = length(ppt))

  ppt_remove <- d %>% 
    select(ppt, enough_sentences) %>% 
    unique() %>%
    filter(!enough_sentences) %>% 
    drop_na() %>% 
    pull(ppt)

  full_set <- d %>% 
    filter(!ppt %in% ppt_remove) %>% 
    count(ppt, topic, genre) %>% 
    summarise(n = n(), .by = ppt) %>% 
    mutate(has_full_set = n == 4)

@@

  result[[1]] <- d %>%
    filter(!is.na(iki), 
           !is_edit,
           ppt %in% ppt_keep,
           !ppt %in% ppt_remove,
           enough_sentences) %>%
    select(ppt, iki, location) %>% 
    drop_na() %>% 
    mutate(too_fast = iki <= 50, 
           too_slow = iki >= 30000,
           across(ppt, ~as.numeric(factor(.)))) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean, 
                          se = se_bin)),
              .by = c(ppt)) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean), 
                     .names = "{.col}"))

  # Sample within each category random data points
  set.seed(365)
  result[[2]] <- d %>% 
    filter(!is.na(iki), 
           !is_edit,
           ppt %in% ppt_keep,
           !ppt %in% ppt_remove,
           enough_sentences,
           iki > 50,
           iki < 30000) %>% 
    mutate(keep = 1:n(),
           keep = sample(keep),
           keep = keep <= n_samples,
           .by = c(ppt, location, genre, topic)) %>% 
   summarise(across(keep, 
             list(mean = mean, 
                  se = se_bin)),
              .by = c(ppt, location)) %>% 
   summarise(across(starts_with("keep"), 
                     list(mean = mean), 
                     .names = "{.col}"), 
              .by = location)

@@

  result[[1]] <- d %>% 
    mutate(too_fast = iki <= 50,
           too_slow = iki >= 30000,
           across(ppt, ~as.numeric(factor(.)))) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean, 
                          se = se_bin)),
              .by = c(ppt)) %>% 
    summarise(across(starts_with("too"), 
                     list(mean = mean), 
                     .names = "{.col}"))

@@

  set.seed(365)
  result[[2]] <- d %>% 
    mutate(keep = 1:n(),
           keep = sample(keep),
           keep = keep <= n_samples,
           .by = c(ppt, location, Lang)) %>% 
   summarise(across(keep, 
             list(mean = mean, 
                  se = se_bin)),
              .by = c(ppt, location)) %>% 
   summarise(across(starts_with("keep"), 
                     list(mean = mean), 
                     .names = "{.col}"), 
              .by = location) %>% 
    mutate(across(location, ~str_replace(., "_", " ")),
           across(location, ~recode(., `sentence before` = "before sentence",
                                  `word before` = "before word")))

  return(result)
}
get_gunnexp2 <- function(file, n_samples){
  d <- read_csv(file) %>%
    filter(!is.na(iki),
           !is.na(isfluent),
           isfluent == 1) %>%
    mutate(ppt = as.numeric(factor(ppt))) %>% 
    select(ppt, iki, xn, location)

  result[[1]] <- d %>% 
  mutate(too_fast = iki <= 50,
         too_slow = iki >= 30000) %>%
  summarise(across(starts_with("too"), 
                   list(mean = mean, 
                        se = se_bin)),
            .by = c(ppt)) %>% 
  summarise(across(starts_with("too"), 
                   list(mean = mean), 
                   .names = "{.col}"))
  # Sample within each category random data points
  set.seed(365)
  result[[2]] <- d %>% 
  filter(iki > 50,
         iki < 30000) %>% 
  mutate(keep = 1:n(),
         keep = sample(keep),
         keep = keep <= n_samples,
         .by = c(ppt, location, xn)) %>% 
  summarise(across(keep, 
                   list(mean = mean, 
                        se = se_bin)),
            .by = c(ppt, location)) %>% 
  summarise(across(starts_with("keep"), 
                   list(mean = mean), 
                   .names = "{.col}"), 
            .by = location)
  return(result)
}
```

```{r}
# Get info about data reduction
n_samples <- 100
c2l1 <- get_c2l1("../data/c2l1.csv", n_samples = n_samples)
cato <- get_cato("../data/cato.csv", n_samples = n_samples)
lift <- get_lift("../data/lift.csv", n_samples = 50, n_ppts = 100)
plantra <- get_plantra("../data/plantra.csv", n_samples = n_samples)
spl2 <- get_spl2("../data/spl2.csv", n_samples = n_samples) 
gunnexp2 <- get_gunnexp2("../data/gunnexp2.csv", n_samples = n_samples)

#lift[[3]]
#lift[[4]]
#plantra[[3]]
#plantra[[4]]
#spl2[[3]] 
```

For all datasets we only used transitions that were not followed by an editing operation.

We removed participants that did not complete all conditions in studies with within-participant factors (reducing the number of participants to 343 in the LIFT data set, and 41 participants in the PLanTra data set). We removed participants that produced less than 10 sentences (LIFT: 109 participants; PLanTra: 3 participants; SPL2: 1 participant)

We removed keystroke intervals that are extremely short ($\le$ 50 msecs) or extremely long ($\ge$ 30 secs). The percentage of remove keystroke data can be found in Table \@ref(tab:datareduction).

From the remaining data we randomly sampled `r n_samples` observations per participant, per condition, and per transition location, with the exception of the LIFT data set. This was done for computational reasons to reduce the time the Bayesian models need to complete. For the LIFT data set we reduced the number of participants to 100 which is substantially more than most of the other data sets in our analysis. Because we included the large number of writing tasks in the LIFT data set as fixed effect, we sampled 50 observations per condition, location and participant. The percentage of keystroke data that went into the final analysis can be found, by transition location, in Table \@ref(tab:datareduction).

```{r datareduction, results='asis'}
c2l1$ds <- "c2l1"
cato$ds <- "cato"
lift$ds <- "lift"
plantra$ds <- "plantra"
spl2$ds <- "spl2"
gunnexp2$ds <- "gunnexp2"

extrem_values <- map_dfr(list(c2l1, cato, lift, plantra, spl2, gunnexp2), ~.x[[1]] %>% 
          mutate(ds = .x$ds)) %>% 
  mutate(across(where(is.numeric), ~round(.*100, 2)),
         across(starts_with("too"), ~str_c(., "%"))) %>% 
  unite("<50ms", starts_with("too_fast"), sep = " (") %>% 
  unite(">30,000ms", starts_with("too_slow"), sep = " (") %>% 
  mutate(across(c(`<50ms`, `>30,000ms`), ~str_c(., ")"))) %>% 
  rename(`$\\le$ 50 msecs` = `<50ms`,
         `$\\ge$ 30 secs` = `>30,000ms`)

random_sample <- 
  map_dfr(list(c2l1, cato, lift, plantra, spl2, gunnexp2), ~.x[[2]] %>% 
          mutate(ds = .x$ds)) %>% 
    mutate(across(where(is.numeric), ~round(.*100, 1)),
           across(starts_with("keep"), ~str_c(., "%"))) %>%
    unite("keep", starts_with("keep"), sep = " (") %>% 
    mutate(across(c(keep), ~str_c(., ")"))) %>%
  pivot_wider(names_from = location, values_from = keep)

datareduction <- left_join(extrem_values, random_sample, by = "ds") %>% 
  relocate(ds) %>% 
  mutate(across(ds, recode, 
                "cato" = "CATO",
                "spl2" = "SPL2", 
                "spl2_shift" = "SPL2 (shift + C)", 
                "plantra" = "PLanTra",
                "lift" = "LIFT",
                "c2l1" = "C2L1",
                "gunnexp2" = "GUNNEXP2")) %>% 
  rename(`Dataset` = ds)

kable(datareduction, 
      caption = "Data reduction. Mean percentage of extreme data removed and the mean percentage of randomly sampled data by transition locattion. Standard error is shown in parentheses.",
        align =c("l", rep("r", 14))) %>%
  kable_styling(position = "center", 
                full_width = T, 
                font_size = 11) %>%
  add_header_above(c(" " = 1, 
                     "Extreme values" = 2,
                     "Randomly sampled data" = 3)) %>%
  column_spec(1, width = "8em")

@@

# Model comparison with simulated data

A general concern with mixture models is that in principle, as the mixture model has more parameters it might simply always lead to a better fit, even though cross-validation is addressing potential problems with overfitting models.

To address this concern we simulated two sets of data. Both data sets have two conditions and 1,000 observations each. The first set of data has as underlying data generating process a mixture model with two mixture components similar to the process described above. The difference between the two conditions is that the mixing proportion is larger for condition 2 than for condition 1, hence long observations are more likely in condition 2. This model can be summarised as followed:

$$
\text{y}_i \sim \theta_\text{condition[i]} \cdot \text{logN}(\mu_1, \sigma^2_1) +\\
(1 - \theta_\text{condition[i]}) \cdot \text{logN}(\mu_2, \sigma^2_2)
$$

The second data set was generated with an unequal variance unimodal model as data generating process. Condition 2 has a larger mean and standard deviation than condition 1. The model can be summaried as followed:

$$
\text{y}_i \sim \text{logN}(\mu_\text{condition[i]}, \sigma^2_\text{condition[i]})
$$

The true parameter values used for each of the two data simulations can be found in Table \@ref(tab:simparam). The simulated data are visualised in Figure \@ref(fig:simdata). The data are simulated to be similarly distributed to keystroke transitions.

```{r simdata, fig.cap= "Data simulated with a bimodal process (left) and a unimodal process (right).", fig.height=4}
files <- list.files("../mixture-model-sim/data", full.names = T)

map_dfr(files, ~read_csv(.) %>% 
          mutate(Dataset = str_remove_all(.x, "../mixture-model-sim/data|/|.csv" ))) %>%
  mutate(across(condition, factor),
         across(condition, ~str_c("Condition: ", condition)),
         across(Dataset, ~recode(., mogdata = "bimodal", uvlmdata = "unimodal"))) %>% 
  ggplot(aes(x = value, colour = condition, fill = condition)) +
  geom_density(alpha = .25) +
  facet_grid(~Dataset, labeller = label_both) +
  scale_colour_colorblind() + 
  scale_fill_colorblind() +
  labs(colour = "", fill = "", x = "Simulated data", y = "Density")

We run 4 models: 2 mixture models, one on the data generated with a mixture process and one on the data generated with the unimodal unequal variance process. We repeated the same using an unimodal unequal variance model. Models were run with 3 chains, with each 6,000 iterations of which 3,000 were warmup. The Stan models uncovered the model parameters of their respective data sets successfully, as shown in Table \@ref(tab:simparam), not less so when the model was applied to data generated with the other underlying process.

```{r}
mog1 <- read_csv("../mixture-model-sim/stanout/mog_mogdata.csv") %>% 
  mutate(data = "mog")
mog2 <- read_csv("../mixture-model-sim/stanout/mog_uvlmdata.csv") %>% 
  mutate(data = "uvlm")
uvlm1 <- read_csv("../mixture-model-sim/stanout/uvlm_uvlmdata.csv") %>% 
  mutate(data = "uvlm")
uvlm2 <- read_csv("../mixture-model-sim/stanout/uvlm_mogdata.csv") %>% 
  mutate(data = "mog")

mog_sim <- bind_rows(mog1, mog2) %>% 
  summarise(across(value, list(est = mean,
                               lo = lower,
                               up = upper),
                   .names = "{.fn}"),
            .by = c(name, data)) %>% 
  mutate(across(where(is.numeric), round, 2),
         across(where(is.numeric), ~ifelse(str_detect(name, "prob"), dezero(., 2), .)),
         across(est, ~str_c(., " [", lo, ", ", up, "]")),
         across(name, ~str_c("\\", .)),
         across(name, ~str_replace(., "a_e", "a^2_1")),
         across(name, ~str_replace(., "ap_e", "a^2_2")),
         across(name, ~str_replace(., "\\[1\\]", "_\\\\text{condition=1}")),
         across(name, ~str_replace(., "\\[2\\]", "_\\\\text{condition=2}"))) %>%
  select(-lo, -up) %>% 
  pivot_wider(names_from = data, values_from = est) %>% 
  mutate(true = c(5, 1, .1, .4, .25, .5),
         across(true, ~ifelse(str_detect(name, "prob"), dezero(., 2), .)),
         across(name, ~str_replace(., "prob", "theta"))) %>% 
  relocate(name, true)

uvlm_sim <- bind_rows(uvlm1, uvlm2) %>% 
  summarise(across(value, list(est = mean,
                               lo = lower,
                               up = upper),
                   .names = "{.fn}"),
            .by = c(name, data)) %>% 
  mutate(across(where(is.numeric), round, 2),
         across(est, ~str_c(., " [", lo, ", ", up, "]")),
         across(name, ~str_c("\\", .)),
         across(name, ~str_replace(., "a_e", "a^2_1")),
         across(name, ~str_replace(., "ap_e", "a^2_2")),
         across(name, ~str_replace(., "\\[1\\]", "_\\\\text{condition=1}")),
         across(name, ~str_replace(., "\\[2\\]", "_\\\\text{condition=2}"))) %>% 
  select(-lo, -up) %>%
  pivot_wider(names_from = data, values_from = est) %>% 
  mutate(true = as.character(c(5, 6, .25, .5))) %>% 
  relocate(name, true)

```{r simparam, results='asis'}
table <- bind_rows(
  mog_sim %>% mutate(model = "mog"),
  uvlm_sim %>% mutate(model = "uvlm")) %>% 
  relocate(model) %>% 
  mutate(across(name, ~str_c("$", ., "$")))

names(table) <- c("Model", "Parameter", "True value", "Bimodal data", "Unimodal data")

kable(table[,-1], 
      caption = "Uncovered parameter estimates with 95% probability interval (PI) and true parameter values for each simulated data set and by model and their respective parameters.",
      align =c("l", rep("r", 3))) %>%
  kable_styling(position = "center", 
                full_width = F, 
                font_size = 11) %>% 
  add_header_above(c(" " = 1, 
                     " " = 1,
                     "Estimate with 95\\% PI" = 2)) %>%
  pack_rows("Model: Bimodal mixture model", 1, 6) %>% 
  pack_rows("Model: Unimodal process", 7, 10)

We used LOO-CV to compare the fit of the two models for each data set. The model comparisons can be found for each data generating process in Table \ref@(tab:loossim). The results show that the mixture model does not always lead to higher predictive performance. Indeed, the mixture model showed a lower predictive performance for the data that were generated with a unimodal process. However, for the data generated with a bimodal process, the mixture model model shows a higher predictive performance. In fact, the ratio of $\Delta\widehat{elpd}$ and its SE, as metric for the strength of evidence, shows that the mixture model performs 11 times better than the unimodal model for the data generated with a bimodal process. In comparison, for the unimodal data, the unimodal model shows only 3 times better than the bimodal mixture model. Thus, even though the mixture model does not necessarily perform better for non-bimodal data but it also doesn't necessarily perform much worse. This contrast is likely a reflection of the increased number of parameters in the mixture model.

```{r loossim, results='asis'}
table <- read_csv("../mixture-model-sim/stanout/modelcomparison.csv") %>% 
  select(1:6) %>% 
  mutate(
#    elpd_se_ratio = abs(elpd_diff / se_diff),
    across(where(is.numeric), round, 0),
    across(where(is.numeric), format, big.mark = ","),
    across(everything(), as.character),
    across(everything(), str_trim),
    across(everything(), str_replace_all, "NaN", "--")) %>% 
  mutate(across(elpd_diff, str_c, " (", se_diff, ")"),
         across(elpd_loo, str_c, " (", se_elpd_loo, ")"),
         across(elpd_diff, str_replace_all, "0 \\( 0\\)", "--"),
         across(c(data, model), ~recode(., mog = "Bimodal mixture model",
                                        uvlm = "Unimodal unequal-variance model"))) %>% 
  select(data, model, starts_with("elpd"))

names(table) <- c("Data", "Model", "$\\Delta\\widehat{elpd}$", "$\\widehat{elpd}$")

kable(table[,-1], 
      caption = "Model comparisons by data set. The top row shows the models with the highest predictive performance. Standard error is shown in parentheses.",
        align =c("l", rep("r", 2))) %>%
  kable_styling(position = "center", 
                full_width = F, 
                font_size = 11) %>%
  footnote("$\\widehat{elpd}$ = predictive performance indicated as expected log pointwise predictive density; $\\Delta\\widehat{elpd}$ = difference in predictive performance relative to the model with the highest predictive performance in the top row.") %>% 
  pack_rows("Data: Bimodal mixture process", 1, 2) %>% 
  pack_rows("Data: Unimodal process", 3, 4)

## Posterior parameter estimates of mixture model

### GUNNEXP2

#### Unconstrained mixture model

##### Posterior parameter estimates

@@

##### Masking effect

@@

#### Constrained mixture model

##### Posterior parameter estimates

@@

### C2L1

@@

### CATO

@@

### PLanTra

@@

### LIFT

@@

### SPL2

@@

## Fit to data

```{r prediction, fig.cap="Fit to data. Comparison of 100 simulated (predicted) sets of data to observed data illustated by model. For illustration the x-axis was truncated at 2,000 msecs.", eval = T, fig.height=10}

run <- FALSE

if(run){
  file <- list.files("../stanout", pattern = "sims", full.names = T, recursive = T)
  ds <- str_remove_all(file, "\\.|\\/|stanout|all_sims.csv") %>% 
    str_to_upper() %>% 
    str_replace(., "PLANTRA", "PLanTra")
  ds <- ds[-6]
  plots <- list()
  plots <- map(1:length(ds), ~get_pdens_plot(file[.], max_iki = 2000, ds = ds[.])) 
  #plots <- map(1, ~get_pdens_plot(file[.], max_iki = 2000, ds = ds[.])) 
  plots_fin <- plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]] / plots[[6]] +
    plot_layout(guides = "collect") &
  #  plot_annotation(tag_levels = 'A') & 
    theme(legend.position = 'bottom')
  saveRDS(plots_fin, "objects/fitplots.rda", compress = "xz")
}

readRDS("objects/fitplots.rda")
```
